### 2021-11-11 NSF2021/Description/specificaim1.tex
figures in preliminary work

\section{Proposed Research}
\label{sec:proposed-work}
%%
\begin{center}
  \begin{figure}
      \includegraphics[width=1\textwidth]{figures/SpecificAim1/Domain.pdf}
  \end{figure}
\end{center}

@@

@@

\subsubsection{Janus-particle self-assembly} 
\begin{wrapfigure}[14]{l}{0.5\textwidth}
\includegraphics[width=0.5\textwidth]{figures/SpecificAim1/multilamelar.pdf}
\caption{\label{fig:multilamelar}
  Experimentalists can transform unilamellar vesicles to multilamellar
  vesicles by altering the quality of the solvent
    \cite{deMeetal21}.
  In \eqref{eq:SL}, the solvent quality enters through the potential
  $f(u)$ and boundary conditions. 
}
\end{wrapfigure}
A number of morphologies arise from the theory. In the linear response case,
take $f(u) = u^2$ so that water structure has its lowest potential value
at $u = 0$, representing bulk water. The solutions of \eqref{eq:SL} have
a boundary layer structure and decay to 0 in the bulk with the decay
length $\rho$.

Consider a polar boundary condition where
$g_i(\mathbf{x}) = (\pi r)^{-1/2} \cos (\theta_i(\mathbf{x}))$
where $\theta_i$ is the angle formed by $\mathbf{x}$, the particle
center $\mathbf{a}_i$, and the particle director $\mathbf{d}_i$.
The normalization $(\pi r)^{-1/2}$ here and others below
provide for a uniform surface energy $\int_{\Gamma_i} g_i^2 \,ds = 1$
when the particles are disks of radius $r$.
The boundary condition is odd along the particle axis 
making the particle head repel the tail.
We simulate the dynamics of 100 particles with random initial position
and orientation. The particles rapidly form chains with their
directors perpendicular to the length of the chain. The equilibrium
structure consists particles positioned on a square grid with alternating
orientations.
\begin{center}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_A1.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_A2.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_A5.png}\\
    Figure 2. Self assembly for a polar boundary condition
    \label{fig:self-assemblyA}
\end{center}

Amphiphilic particles have a hydrophobic tail and a hydrophilic head
and are accounted for as follows.
The hydrophilic side takes the value $u = 0$ since the
apolar headgroup of a lipid, for example, does not alter the structure
of adjacent waters. The hydrophobic side takes the value $u > 0$. 
The interaction between particles is attractive. Using
$(3\pi )^{-1/2}(1 + \cos(\theta_i(\mathbf{x})))$
as a simple modification to the previous boundary condition 
produces two-dimensional micelles and bilayers \cite{Fu2018_SIAM}.
Researchers have asked us if the phenomenon is robust. To show
that it is, we use 
\begin{equation}
\label{eq:vonMises}
g_i(\mathbf{x}) =
\sqrt{
  \frac{\exp( \kappa_i(\theta_i(\mathbf{x}) - \theta^0_i))}
  {2\pi r_i I_0(\kappa_i)}}
\end{equation}
instead where the radicand is the von Mises 
distribution with director angle $\theta^0_i$ and Bessel function
$I_0$ of order zero. The square of this function behaves as a normal
distribution with variance like $1/\kappa_i$. To replicate
experimental conditions, the concentration $\kappa_i$ and radii
$r_i$ are randomized. Using the same initial configuration, the
particles form small groups. The small groups combine to form bilayers.
\begin{center}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_B1.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_B2.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_B4.png}\\
    Figure 3. Self assembly for a amphiphilic boundary condition
    \label{fig:self-assemblyA}
\end{center}

Multilamelar bilayers arise with an increase by shifting the boundary condition
further. Let $g_i(\mathbf{x}) = (9\pi )^{-1/2}(2 + \cos(\theta_i(\mathbf{x})))$.
This gives a hydrophobic particle with greater intensity of hydrophobicity
at $\theta_i = 0$ than at $\theta_i = \pi$.  The initial self-assembly
is similar to that of the amphiphilic particle case, except that the bilayers
do not stay well separated.  Rather, the bilayers stack on top of each
other as a consequence of the interfacial tension of exposed particle heads. 
\begin{center}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_C1.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_C2.png}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100_C4.png}\\
    Figure 3. Self assembly for hydrophobic particles with bias.
    \label{fig:self-assemblyA}
\end{center}

In summary, the hydrophibic interaction-mobility problem model yields
rich morphologies that are valuable in the synthesizing multiphasic and
anisotropic colloids. The goal is to obtain numerically aided 
syntheses and assembly of soft matter materials.
Applications in this area include cargo-carrying capacity, 
catalytic activity, or stimuli-responsive shape-change to amplify colloid
utility for a wide range of  spanning environmental remediation to drug
delivery. The main experimental challenges include the ability to
decouple particle surface properties.
The main modeling challenge is to translate the phenomenological
parameters, e.g. boundary conditions, into measurable quantities
such as contact angles. Collaboration with experimental groups will
use numerical methods to predict parameter regimes 
to control particle-particle interactions. In Specific Aim 3, we
discuss how to include ions and surface charge.

@@

Here, the order parameter $u$ is the mean number of nearest
neighbors per water molecule.
The value $u_0 = 4$ gives a local minimum for the ideal
tetrahadral water network found at hydrophobic surfaces.  
The value $u_1 \sim 4.4$ gives the absolute minimum
in the bulk where there is a mixture of four and five-coordinated
waters. The equation \eqref{eq:SL} takes the form of a
steady-state Allen-Cahn equation. 
When particles are well-separated, the energetic cost for filling
the interstitial region with $u_0$-phase water is prohibitive
and there is a phase shift from the $u_0$ to $u_1$ water order
---the attraction landscape is more
or less identical to the linear response theory. Conversely, when
particles are proximal, it becomes energetically favorable
to coalesce $u_0$-phases. This introduces a line tension
for the phase transition.

>>>>

\section{Proposed Research}
\label{sec:proposed-work}
%%

@@

  \begin{figure}[h!]
\begin{center}
  \includegraphics[width=0.8\textwidth]{figures/SpecificAim1/Domain.pdf}
\end{center}  
  \caption{\label{fig:flow_map} A collection of 3D
   amphiphilic particles suspended in a solvent. The solvent velocity
   satisfies the Stokes equations, and the hydrophobic forces depend on
   the solvent activity that satisfies the screened Laplace equation.
   The geometry is updated by solving the mobility problem.}     
  \end{figure}

@@

\subsubsection{Janus-particle self-assembly} 
In the linear response case,
take $f(u) = u^2$ so that water structure has its lowest potential value
at $u = 0$, representing bulk water. The solutions of \eqref{eq:SL} have
a boundary layer structure and decay to 0 in the bulk with the decay
length $\rho$. Nevertheless, a number of morphologies arise by altering
boundary conditions.

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100A1.pdf}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100A2.pdf}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100A3.pdf}
    \end{center}
  \caption{
    Self assembly with a polar boundary condition leads to checkboard pattern.
    Green is for $u < 0$, yellow is for $u > 0$, and blue is for $u = 0$.
    \label{fig:self-assemblyA}}
\end{figure}

Consider a polar boundary condition where
$g_i(\mathbf{x}) = (\pi r)^{-1/2} \cos (\theta_i(\mathbf{x}))$
where $\theta_i$ is the angle formed by $\mathbf{x}$, the particle
center $\mathbf{a}_i$, and the particle director $\mathbf{d}_i$.
The normalization $(\pi r)^{-1/2}$ here and others below
provide for a uniform surface energy $\int_{\Gamma_i} g_i^2 \,ds = 1$
when the particles are disks of radius $r$.
The boundary condition is odd along the particle axis 
making the particle head repel the tail.
We simulate the dynamics of 100 particles with random initial position
and orientation. The particles rapidly form chains with their
directors perpendicular to the length of the chain. The equilibrium
structure consists particles positioned on a square grid
in a checkerboard pattern. This patterning is a consequence of energy
minimality and allows each particle to simultaneously
coordinate its head with that of three other particles
and its tail with that of three other particles.

\begin{figure}[h!]
\begin{center}  
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100B1.pdf}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100B2.pdf}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100B3.pdf}\\
\end{center}    
\caption{With an amphiphilic boundary condition particles self-assemble
  into bilayers. The patterning shields the hydrophobic core (yellow) from the
  aqueous phase. 
      \label{fig:self-assemblyA}}
\end{figure}
Amphiphilic particles have a hydrophobic tail and a hydrophilic head
and are accounted for as follows.
The hydrophilic side takes the value $u = 0$ since the
apolar headgroup of a lipid, for example, does not alter the structure
of adjacent waters. The hydrophobic side takes the value $u > 0$. 
The interaction between particles is attractive. Using
$(3\pi )^{-1/2}(1 + \cos(\theta_i(\mathbf{x})))$
as a simple modification to the previous boundary condition 
produces two-dimensional micelles and bilayers \cite{Fu2018_SIAM}.
Researchers have asked us if the phenomenon is robust. To show
that it is, we use 
\begin{equation}
\label{eq:vonMises}
g_i(\mathbf{x}) =
\sqrt{
  \frac{\exp( \kappa_i(\theta_i(\mathbf{x}) - \theta^0_i))}
  {2\pi r_i I_0(\kappa_i)}}
\end{equation}
instead where the radicand is the von Mises 
distribution with director angle $\theta^0_i$ and Bessel function
$I_0$ of order zero. The square of this function behaves as a normal
distribution with variance like $1/\kappa_i$. To replicate
experimental conditions, the concentration $\kappa_i$ and radii
$r_i$ are randomized. Using the same initial configuration, the
particles form small groups. The small groups combine to form bilayers.

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100C1.pdf}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100C2.pdf}
    \includegraphics[width=0.32\textwidth]{figures/SpecificAim1/N100C3.pdf}\\
    \caption{This figure is for hydrophobic particles with a bias. The
      hydrophobic intensity is greater on one side of the particle (magenta)
      than the other (yellow).
    \label{fig:self-assemblyA}}
\end{center}
\end{figure}

Multilamelar bilayers arise with an increase by shifting the boundary condition
further. Let $g_i(\mathbf{x}) = (9\pi )^{-1/2}(2 + \cos(\theta_i(\mathbf{x})))$.
This gives a hydrophobic particle with greater intensity of hydrophobicity
at $\theta_i = 0$ than at $\theta_i = \pi$.  The initial self-assembly
is similar to that of the amphiphilic particle case, except that the bilayers
do not stay well separated.  Rather, the bilayers stack on top of each
other as a consequence of the interfacial tension of exposed particle heads \cite{deMeetal21}.

In summary, the hydrophibic interaction-mobility problem model yields
rich morphologies that are valuable in the synthesizing multiphasic and
anisotropic colloids \cite{Bradley2016,Mallory2017,Bradley2017}.
The main experimental challenges include the ability to
decouple particle surface properties.
The goal is to obtain numerically aided 
syntheses and assembly of soft matter materials.
Applications in this area include cargo-carrying capacity, 
catalytic activity, or stimuli-responsive shape-change to amplify colloid
utility for a wide range of  spanning environmental remediation to drug
delivery \cite{McBr21, HaBr20}.

%\begin{wrapfigure}[14]{l}{0.5\textwidth}
%\includegraphics[width=0.5\textwidth]{figures/SpecificAi%m1/multilamelar.pdf}
%\caption{\label{fig:multilamelar}
%  Experimentalists can transform unilamellar vesicles to% multilamellar
%  vesicles by altering the quality of the solvent
%    \cite{deMeetal21}.
%  In \eqref{eq:SL}, the solvent quality enters through t%he potential
%  $f(u)$ and boundary conditions. 
%}
%\end{wrapfigure}
The main modeling challenge is to translate the phenomenological
parameters, e.g. boundary conditions, into measurable quantities
such as contact angles. Collaboration with experimental groups will
use numerical methods to predict parameter regimes 
to control particle-particle interactions. In Specific Aim 3, we
discuss how to include ions and surface charge.

@@

Here, the order parameter $u$ is the mean number of nearest
neighbors per water molecule.
The value $u_0 = 4$ gives a local minimum for the ideal
tetrahadral water network found at hydrophobic surfaces.  
The value $u_1 \sim 4.4$ gives the absolute minimum
in the bulk where there is a mixture of four and five-coordinated
waters.
When particles are well-separated, the energetic cost for filling
the interstitial region with $u_0$-phase water is prohibitive
and there is a phase shift from the $u_0$ to $u_1$ water order
---the attraction landscape is more
or less identical to the linear response theory. Conversely, when
particles are proximal, it becomes energetically favorable
to coalesce $u_0$-phases. This introduces a line tension
for the phase transition.

The equation \eqref{eq:SL} takes the form of a steady-state
Allen-Cahn equation. The Allen-Cahn equation is used in phase field
formulations of immisible fluids, for example.
PI Ryham and colaborators launched the simulation and asymptotic analysis
of phase field functionals of membrane elastic energy,
devising functionals for  
for the squared mean curvature (Willmore) energy \cite{0951-7715-18-3-016}, 
spontaneous curvature \cite{Du05} and Gaussian curvature energy \cite{DuEuler},
as well as energetic variational approaches to coupling membrane elasticity to fluids \cite{QiangDu09}.

### 2021-11-11 NSF2021/Description/specificaim2.tex
figures in preliminary work

\begin{wrapfigure}[]{l}{1.8in}
  \centering
  \input{figures/doublewell.tikz}
  \caption{\label{fig:well} $U_1 = 0.5$, $U_2 = 2$, $a = 1$, $b = 0.4$.}
\end{wrapfigure}
In the PIs collaboration, they have developed numerical methods to
solve~\eqref{eq:SL} when $f(u) = u^2$. With this choice of $f(u)$, the
resulting PDE is the linear screened Laplace equation which is
efficiently solved with a double-layer potential representation. For a
more general $f$, such as a double well potential, we will interpret the
solution of~\eqref{eq:SL} as a root finding problem for the operator
\begin{align}
  \label{eq:F}
  F[u] = -\rho^2 \Delta u + f'(u).
\end{align}
This non-linear PDE will be solved with Newton's method
\begin{align*}
  u^{N+1} = u^{N} - \alpha dF^{-1} F(u^N),
\end{align*}
where $dF = -\rho^2 \Delta + f''(u^N)$ is the Fr\'{e}chet derivative of
$F$, and $\alpha$ is the step size which will be determined using
backtracking. As expected, the biggest challenge amounts to solving
\begin{align}
  \label{eq:NewtonUpdate}
  - \rho^2 \Delta v + f''(u^{N}) v = F[u^N].
\end{align}
For a general double well potential $f(u)$, this is a variable
coefficient elliptic PDE, and developing an integral equation
formulation is challenging. We propose to instead use an inexact
Newton's method where $dF^{-1}$ is defined in terms of the piecewise
quadratic approximation
\begin{align*}
  f(u) \approx \tilde{f}(u) = bu + \begin{cases}
    a(u - U_1)^2, &\mbox{if } u \leq (U_0 + U_1)/2 \\
    (u - U_2)^2, &\mbox{if } u > (U_0 + U_1)/2.
  \end{cases}
\end{align*}
In this case, the update step of Newton's method~\eqref{eq:NewtonUpdate}
amounts to solving the PDE
\begin{align}
  \label{eq:screenedPoisson}
  -\rho^2 \Delta v + \tilde{f}''(u^{N})v = F[u^N],
\end{align}
\begin{wrapfigure}[]{r}{1.8in}
  \centering
  \input{figures/residuals.tikz}
  \caption{\label{fig:residuals} $U_1 = 0.5$, $U_2 = 2$, $a = 1$, $b =
  0.4$.}
\end{wrapfigure}
which is a screened Poisson equation with piecewise a constant screening
coefficient. We test the efficacy of Newton's method applied
to~\eqref{eq:F} by considering its one-dimensional version in $(0,1)$
with the Dirichlet boundary conditions $u(0) = U_0$ and $u(1) = U_1$.
These parameters result in the double well potential illustrated in
Figure~\ref{fig:well}(a), and the inexact Newton's method converges to
the solution in Figure~\ref{fig:well}(b). Figure~\ref{fig:residuals}
shows the max norm of the residual at each Newton iterate when using
the exact second derivative (red curve) and the piecewise constant
approximate (blue curve).

>>>>

In the PIs collaboration, they have developed numerical methods to
solve~\eqref{eq:SL} when $f(u) = u^2$. With this choice of $f(u)$, the
resulting PDE is the linear screened Laplace equation which is
efficiently solved with a double-layer potential representation. For a
more general $f$, such as a double well potential, we will interpret the
solution of~\eqref{eq:SL} as a root finding problem for the operator
\begin{align}
  \label{eq:F}
  F[u] = -\rho^2 \Delta u + f'(u).
\end{align}
%\begin{wrapfigure}[13]{l}{1.8in}
%  \centering
%  \input{figures/doublewell.tikz}
%  \caption{\label{fig:well} $U_1 = 0.5$, $U_2 = 2$, $a = 1$, $b = 0.4$.}
%\end{wrapfigure}
This non-linear PDE will be solved with Newton's method
\begin{align*}
  u^{N+1} = u^{N} - \alpha dF^{-1} F(u^N),
\end{align*}
where $dF = -\rho^2 \Delta + f''(u^N)$ is the Fr\'{e}chet derivative of
$F$, and $\alpha$ is the step size which will be determined using
backtracking.
As expected, the biggest challenge amounts to solving
\begin{align}
  \label{eq:NewtonUpdate}
  - \rho^2 \Delta v + f''(u^{N}) v = F[u^N].
\end{align}
\begin{wrapfigure}[20]{r}{1.8in}
  \centering
  \input{figures/doublewell.tikz}  \\
  \input{figures/residuals.tikz}
  \caption{\label{fig:residuals} A double well (top) and the error (bottom)
    for the case $U_1 = 0.5$, $U_2 = 2$, $a = 1$, $b = 0.4$.}
\end{wrapfigure}
For a general double well potential $f(u)$, this is a variable
coefficient elliptic PDE, and developing an integral equation
formulation is challenging. We propose to instead use an inexact
Newton's method where $dF^{-1}$ is defined in terms of the piecewise
quadratic approximation
\begin{align*}
  f(u) \approx \tilde{f}(u) = bu + \begin{cases}
    a(u - U_1)^2, &\mbox{if } u \leq (U_0 + U_1)/2 \\
    (u - U_2)^2, &\mbox{if } u > (U_0 + U_1)/2.
  \end{cases}
\end{align*}
In this case, the update step of Newton's method~\eqref{eq:NewtonUpdate}
amounts to solving the PDE
\begin{align}
  \label{eq:screenedPoisson}
  -\rho^2 \Delta v + \tilde{f}''(u^{N})v = F[u^N],
\end{align}
which is a screened Poisson equation with piecewise a constant screening
coefficient. We test the efficacy of Newton's method applied
to~\eqref{eq:F} by considering its one-dimensional version in $(0,1)$
with the Dirichlet boundary conditions $u(0) = U_0$ and $u(1) = U_1$.
These parameters result in the double well potential illustrated in
Figure~\ref{fig:well}(a), and the inexact Newton's method converges to
the solution in Figure~\ref{fig:well}(b). Figure~\ref{fig:residuals}
shows the max norm of the residual at each Newton iterate when using
the exact second derivative (red curve) and the piecewise constant
approximate (blue curve).
